### 2013-10-14 pubform.tex

%%
%% This is file `pubform.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% acmconf.dtx  (with options: `pubform')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from pubform.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file acmconf.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% to test the checksum, uncomment    \OnlyDescription
%% in the driver
%%\fi
%%
%% \CharacterTable
%%  {Upper-case    \A\B\C\D\E\F\G\H\I\J\K\L\M\N\O\P\Q\R\S\T\U\V\W\X\Y\Z
%%   Lower-case    \a\b\c\d\e\f\g\h\i\j\k\l\m\n\o\p\q\r\s\t\u\v\w\x\y\z
%%   Digits        \0\1\2\3\4\5\6\7\8\9
%%   Exclamation   \!     Double quote  \"     Hash (number) \#
%%   Dollar        \$     Percent       \%     Ampersand     \&
%%   Acute accent  \'     Left paren    \(     Right paren   \)
%%   Asterisk      \*     Plus          \+     Comma         \,
%%   Minus         \-     Point         \.     Solidus       \/
%%   Colon         \:     Semicolon     \;     Less than     \<
%%   Equals        \=     Greater than  \>     Question mark \?
%%   Commercial at \@     Left bracket  \[     Backslash     \\
%%   Right bracket \]     Circumflex    \^     Underscore    \_
%%   Grave accent  \`     Left brace    \{     Vertical bar  \|
%%   Right brace   \}     Tilde         \~}

\documentclass[submit]{acmconf}
\usepackage{flushend}

\CopyrightText{LEAVE BLANK THE LAST 3.81 cm
               (1.5'') OF THE LEFT COLUMN ON THE FIRST PAGE
               FOR THE COPYRIGHT NOTICE }
\ConferenceName{1. Conference on Designing a \LaTeX2e Class for
                Typesetting ACM Papers, Hawaii 2000}
\ConferenceShortName{CONF-2000}

\begin{document}
\date{31. December 1999}
\title{ACM SIG Proceedings Format}
\author{\Author{1st Author}                               \\
        \Address{First author's affiliation}              \\
        \Address{1st line of address}                     \\
        \Address{2st line of address}                     \\
        \Address{Telephone number, incl.\ country code}   \\
        \Email{1st author's email address}                \\
        \and
        \Author{2nd Author}                               \\
        \Address{First author's affiliation}              \\
        \Address{1st line of address}                     \\
        \Address{2st line of address}                     \\
        \Address{Telephone number, incl.\ country code}   \\
        \Email{2nd author's email address}                \\
        \and
        \Author{3rd Author}                               \\
        \Address{First author's affiliation}              \\
        \Address{1st line of address}                     \\
        \Address{2st line of address}                     \\
        \Address{Telephone number, incl.\ country code}   \\
        \Email{3rd author's email address}                \\
       }
\maketitle

\begin{abstract}
In this paper, we describe the formatting guidelines for ACM SIG
Proceedings.
\end{abstract}

\begin{keywords}
Guides, instructions, authors kit, conference publications.
\end{keywords}

\section{Introduction}

The proceedings are the records of the conference. ACM hopes to
give these conference by-products a single, high-quality appearance.
To do this, we ask that authors follow some simple guidelines. In
essence, we ask you to make your paper look exactly like this
document. The easiest way to do this is simply to down-load a
template from \cite{template.1999}, and replace the content with
your own material.

\section{Page Size}

All material on each page should fit within a rectangle of
18 x 23.5 cm (6" x 9.25"), centered on the page, beginning
2.54 cm (1") from the top of the page and ending with
2.54 cm (1") from the bottom.  The right and left margins
should be 1.9 cm (.75'').  The text should be in two
8.45 cm (3.33") columns with a .83 cm (.33") gutter.

\section{Typeset Text}

Prepare your submissions on a typesetter or word processor.

\subsection{Normal or Body Text}

Please use a 9-point Times Roman font, or other Roman font with
serifs, as close as possible in appearance to Times Roman in
which these guidelines have been set. The goal is to have a
9-point text, as you see here. Please use sans-serif or
non-proportional fonts only for special purposes, such as
distinguishing source code text. The Press 9-point font available
to users of Script is a good substitute for Times Roman. If
Times Roman is not available, try the font named Computer
Modern Roman. On a Macintosh, use the font named Times.
Right margins should be justified, not ragged.

\subsection{Title and Authors}

The title (Helvetica 18-point bold), authors' names (Helvetica
12-point) and affiliations (Helvetica 10-point) run across the full
width of the page - one column wide. We also recommend phone number
(Helvetica 10-point) and e-mail address (Helvetica 12-point). See the
top of this page for three addresses. If only one address is needed,
center all address text. For two addresses, use two centered tabs, and
so on. For more than three authors, you may have to
improvise\footnote{If necessary, you may place some address
information in a foortone, or in a named section at the end of your
paper.}

\subsection{First Page Copyright Notice}

Leave 3.81 cm (1.5") of blank space at the bottom of the left
column of the first page for the copyright notice.

\subsection{Subsequent Pages}

For pages other than the first page, start at the top of the page, and
continue in double-column format.  The two columns on the last page
should be of equal length.

\subsection{References and Citations}

Footnotes should be Times New Roman 9-point.

Use the standard Communications of the ACM format for references -
that is, a numbered list at the end of the article, ordered
alphabetically by first author, and referenced by numbers in
brackets \cite{anderson.1992}. See the examples of citations at
the end of this document
\cite{conger.1995,mackay.1995,schwartz.1995}. Within this template
file, use the style named references for the text of your citation.

References should be published materials accessible to the public.
Internal technical reports may be cited only if they are easily
accessible (i.e. you can give the address to obtain the report within
your citation) and may be obtained by any reader. Proprietary
information may not be cited. Private communications should be
acknowledged, not referenced  (e.g., ``[Robertson, personal
communication]'').

\subsection{Page Numbering, Headers and Footers}

Do not include headers, footers or page numbers in your submission.
These will be added when the publications are assembled.

\section{Sections}

The heading of a section should be in Times New Roman 12-point bold in
all-capitals flush left.  Sections and subsequent subsections should
be numbered and flush left.

\subsection{Subsections}

The heading of subsections should be in Times New Roman 12-point bold
with only the initial letters capitalized. (Note: For subsections and
subsubsections, a word like the or a is not capitalized unless it is
the first word of the header.)

\subsubsection{Subsubsections}

The heading for subsubsections should be in Times New Roman 11-point
italic with initial letters capitalized.

\subsubsubsection{Subsubsubsection}

The heading for subsubsubsections should be in
Times New Roman 11-point italic with initial letters capitalized.

\subsubsubsubsection{Subsubsubsubsection}

The heading for subsubsubsubsections should be in
Times New Roman 11-point italic with initial letters capitalized.

\section{Figures/Captions}

Place Tables/Figures/Images in text as close to the reference as
possible.  It may extend across both columns to a maximum width of
17.78 cm (7'').

Captions should be Times New Roman 9-point bold.  They should be
numbered (e.g., ``Table 1'' or ``Figure 2'') and be centered beneath
each table, figure or image.

\section{Acknowledgments}

Our thanks to ACM SIGCHI for allowing us to modify templates they had
developed.

\medskip

\bibliographystyle{plain}
\bibliography{pubform}

Columns on Last Page Should Be Made Equal Length
\end{document}
\endinput
%%
%% End of file `pubform.tex'.